---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(dplyr)
library(ggplot2)
```

1. Load data
```{r}
activity<-read.csv("activity.csv")
```

2.Change Date from factor to Date
```{r}
activity$date <- as.Date(activity$date, format="%Y-%m-%d")
```


## What is mean total number of steps taken per day?
1. Obtain total number of steps per day
```{r}
total_step<-group_by(activity,date) %>% summarize(steps=sum(steps,na.rm=TRUE))
```

2. Histogram plot
```{r}
g_total<- ggplot(data=total_step,aes(x=steps)) 
g1 <- g_total + theme_bw() +ylim(0,6)
g1+ geom_histogram(fill="blue", binwidth=500) + labs(title="Total Steps Taken Each Day",x="Total Steps/ Date")
```

3. Calculate mean and median of total steps per day
```{r}
mean(total_step$steps)
median(total_step$steps)
```


## What is the average daily activity pattern?
1. Obtain average steps by interval
```{r}
mean_interval <- group_by(activity,interval) %>% summarize(steps=mean(steps,na.rm=TRUE))
```

2. Activity pattern plot
```{r}
g_mean<- ggplot(data=mean_interval,aes(x=interval,y=steps)) 
g2 <- g_mean + geom_line(col="blue") + theme_bw() 
g2+labs(title="Average Steps Taken at 5 min Interval",x="5 minute Interval", y="Average Steps")
```

3. Calculate max interval step
```{r}
mean_interval$interval[which.max(mean_interval$steps)]
```


## Imputing missing values
1. Calculate number of NA value
```{r}
sum(is.na(activity))
```

2. Replace NA value with mean interval value
```{r}
update<-activity
for (i in mean_interval) {
    update[update$interval == i & is.na(update$steps),]$steps <- mean_interval$steps[mean_interval$interval==i]
}
```

3. Updated total number of steps per day
```{r}
update_step<-group_by(update,date) %>% summarize(steps=sum(steps))
```

4. Updated histogram plot
```{r}
g_update<- ggplot(data=update_step,aes(x=steps)) 
g3 <- g_update + theme_bw() 
g3+ geom_histogram(fill="blue", binwidth=500) + labs(title="Updated Total Steps Taken Each Day",x="Total Steps/ Date")
```

## Are there differences in activity patterns between weekdays and weekends?
1. Create column with weekday vs weekend
```{r}
activity$days <- factor(weekdays(activity$date))
activity$weekdays<-""
activity[activity$day == "Saturday" | activity$day == "Sunday", ]$weekdays <- "weekend" 
activity[!(activity$day == "Saturday" | activity$day == "Sunday"), ]$weekdays <- "weekday" 
```

2. Average steps by interval
```{r}
mean_interval <- group_by(activity,weekdays,interval) %>% summarize(steps=mean(steps,na.rm=TRUE))
```

3. Activity pattern plot by weekday vs weekend
```{r}
g_mean<- ggplot(data=mean_interval,aes(x=interval,y=steps)) 
g2 <- g_mean + geom_line(col="blue") + theme_bw() + facet_grid(weekdays~.) 
g2+labs(title="Average Steps Taken at 5 min Interval",x="5 minute Interval", y="Average Steps")
```
